---
title: "alex test"
output: html_notebook
---

## Load Library Package

```{r}
library(tidyverse)
```

## Generate data from example procedure

```
tl_snp$x1 <- tl_snp$rs10808746 * eff$rs10808746
tl_snp$x2 <- tl_snp$rs11556505 * eff$rs11556505
tl_snp$x3 <- tl_snp$rs2075650 *  eff$rs2075650
tl_snp$x4 <- tl_snp$rs2143571 *  eff$rs2143571
tl_snp$x5 <- tl_snp$rs2283228 *  eff$rs2283228
tl_snp$x6 <- tl_snp$rs2896019 *  eff$rs2896019
tl_snp$x7 <- tl_snp$rs429358  *  eff$rs429358
tl_snp$x8 <- tl_snp$rs4402960 *  eff$rs4402960
tl_snp$x9 <- tl_snp$rs6006611 *  eff$rs6006611
tl_snp$x10 <- tl_snp$rs738409 *  eff$rs738409
tl_snp$x11 <- tl_snp$rs7412 *    eff$rs7412
tl_snp$x12 <- tl_snp$rs7901695 * eff$rs7901695
tl_snp$x13 <- tl_snp$rs8181588 * eff$rs8181588
tl_snp$x14 <- tl_snp$rs9331888 * eff$rs9331888
tl_snp$x15 <- tl_snp$rs9331942 * eff$rs9331942
tl_snp$x16 <- tl_snp$rs985421 *  eff$rs985421
tl_snp$x17 <- tl_snp$rs9877502 * eff$rs9877502
```

## Build Reprex data

First, build toy tall table with random data and specific variable names, so that I can replicate the wide table from the original email.

```{r}

var_name <- c("rs10808746", "rs11556505", "rs2075650",
              "rs2143571", "rs2283228", "rs2896019",
              "rs429358", "rs4402960", "rs6006611",
              "rs738409", "rs7412", "rs7901695",
              "rs8181588", "rs9331888", "rs9331942",
              "rs985421", "rs9877502")

x_set <- runif(1700, 0.0, 15.45)

var_val_1 <- seq(length.out = 17, by = 18.6)
var_val_2 <- seq(length.out = 17, by = 33.1)
var_val_3 <- sample(x_set, size = 17)
var_val_4 <- sample(x_set, size = 17)
var_val_5 <- sample(x_set, size = 17)

tbl_snp <- tibble(var_name, var_val_1)
tbl_eff <- tibble(var_name, var_val_2)
tbl_snp3 <- tibble(var_name, var_val_3)
tbl_snp4 <- tibble(var_name, var_val_4)
tbl_snp5 <- tibble(var_name, var_val_5)

tbl_snp_wide <- tbl_snp %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_1)

tbl_snp3 <- tbl_snp3 %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_3)
tbl_snp4 <- tbl_snp4 %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_4)
tbl_snp5 <- tbl_snp5 %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_5)

tbl_eff_wide <- tbl_eff %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_2)

tbl_snp_longer <-  bind_rows(tbl_snp_wide, tbl_snp3, tbl_snp4, tbl_snp5)

tlb_eff_wide2 <- tbl_eff_wide
```


Tranform example data so that it is similar to the problem discussed last night.

- tbl_eff_wide is 1 row and 17 variables
- tbl_snp_longer is 4 rows with the same 17 variable names, plus some dummy starwars data added.  

```{r}
tbl_snp_longer <- bind_cols(head(starwars, 4),  # skip step when replicating on real data
                            tbl_snp_longer)     # skip step when replicating on real data
tbl_eff_wide
tbl_snp_longer
```

## Pivot wide data so that it is truly long. 

The end result is one tibble 65 rows by 15 variables.  The other tibble is 17 rows by 2 columns

```{r}
snp_tall_too <- tbl_snp_longer %>% 
  pivot_longer(starts_with("rs"), 
               names_to = "snp_var_name", 
               values_to = "snp_value")
snp_tall_too

eff_tall_too <- tbl_eff_wide %>% 
  pivot_longer(starts_with("rs"), 
               names_to = "eff_var_name",
               values_to = "eff_value")
eff_tall_too
```

## Join & Mutate

Now Join by the "rs" variable name as the join key.  Then use `mutate` to perform the vecorized math.  Store the new answer in the variable `product`. 


```{r}
foo <- left_join(snp_tall_too, eff_tall_too, by = c("snp_var_name" = "eff_var_name")) %>% 
  mutate(product = snp_value * eff_value) %>% 
  arrange(snp_var_name) %>% 
  select(-films, -starships, -vehicles) %>% 
  select(name, gender, eye_color, snp_var_name, snp_value, eff_value, product, everything())
foo

write_csv(foo, "my_new_data.csv")
```


