---
title:  README
author: "John Little"
date:  "`r Sys.Date()`"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# reprex code 

<!-- badges: start -->
<!-- badges: end -->

### Question

How to calculate variables from to differet tibbles and sum the results

### Answer

1. TRANSFORM: pivot the data from wide to long
1. TRANSFORM: join the the two tables by the variable ID
1. ANALYSIS: calculate the score
1. ANALYSIS: sum the score

See the Rmd for **complete and working code example**:  [reprex_exmple.Rmd](https://github.com/libjohn/code_example_join_and_sum/blob/master/reprex_exmple.Rmd)

## Summary details below...

The following is a [Reprex](https://reprex.tidyverse.org/), i.e. a minimally reproducible example.

### Load Library Package

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

### Generate data from example procedure

```
tl_snp$x1 <- tl_snp$rs10808746 * eff$rs10808746
tl_snp$x2 <- tl_snp$rs11556505 * eff$rs11556505
tl_snp$x3 <- tl_snp$rs2075650 *  eff$rs2075650
tl_snp$x4 <- tl_snp$rs2143571 *  eff$rs2143571
tl_snp$x5 <- tl_snp$rs2283228 *  eff$rs2283228
tl_snp$x6 <- tl_snp$rs2896019 *  eff$rs2896019
tl_snp$x7 <- tl_snp$rs429358  *  eff$rs429358
tl_snp$x8 <- tl_snp$rs4402960 *  eff$rs4402960
tl_snp$x9 <- tl_snp$rs6006611 *  eff$rs6006611
tl_snp$x10 <- tl_snp$rs738409 *  eff$rs738409
tl_snp$x11 <- tl_snp$rs7412 *    eff$rs7412
tl_snp$x12 <- tl_snp$rs7901695 * eff$rs7901695
tl_snp$x13 <- tl_snp$rs8181588 * eff$rs8181588
tl_snp$x14 <- tl_snp$rs9331888 * eff$rs9331888
tl_snp$x15 <- tl_snp$rs9331942 * eff$rs9331942
tl_snp$x16 <- tl_snp$rs985421 *  eff$rs985421
tl_snp$x17 <- tl_snp$rs9877502 * eff$rs9877502
```

### Build Reprex data

First, build toy tall table with random data and specific variable names, so that I can replicate the wide table from the original email.

```{r}

var_name <- c("rs10808746", "rs11556505", "rs2075650",
              "rs2143571", "rs2283228", "rs2896019",
              "rs429358", "rs4402960", "rs6006611",
              "rs738409", "rs7412", "rs7901695",
              "rs8181588", "rs9331888", "rs9331942",
              "rs985421", "rs9877502")

var_val_1 <- seq(length.out = 17, by = 18.6)
var_val_2 <- seq(length.out = 17, by = 33.1)

tbl_snp <- tibble(var_name, var_val_1)
tbl_eff <- tibble(var_name, var_val_2)

tbl_snp_wide <- tbl_snp %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_1)

tbl_eff_wide <- tbl_eff %>% 
  pivot_wider(names_from = var_name,
              values_from = var_val_2)
```

Example data similar to that discussed in the email

```{r}
tbl_snp_wide
tbl_eff_wide
```

#### Pivot longer

["Tall" data](https://en.wikipedia.org/wiki/Wide_and_narrow_data) (a.k.a "long", a.k.a "narrow") is often easier or more convenient for calculations.

```{r}
tbl_snp_tall <- tbl_snp_wide %>% 
  pivot_longer(starts_with("rs"), 
               names_to = "snp_var_name", 
               values_to = "snp_value")
tbl_eff_tall <- tbl_eff_wide %>% 
  pivot_longer(starts_with("rs"), 
               names_to = "eff_var_name",
               values_to = "eff_value")

head(tbl_snp_tall)
head(tbl_eff_tall)
```

### Combine tibbles 

Combine the column values from both tables for later manipulation

#### Easy

Using the `bind_cols()` function.  This is easy to do but each table must be the same size and the data must be in the same order.

```{r}
big_tbl <- bind_cols(tbl_snp_tall, tbl_eff_tall)
head(big_tbl)
```

#### More precise via left_join

Using a join function allows matching by data identifiers, before merging variable values.

There are different types of `join` functions (see [dplyr](https://dplyr.tidyverse.org/reference/join.html).  In this case we want a `left_join()` -- most people want to do a left_join, but it makes sense to verify which type of join you need.


```{r}
join_tbl <- left_join(tbl_snp_tall, tbl_eff_tall, by = c("snp_var_name" = "eff_var_name"))

head(join_tbl)
```

### Calculate new variable

score = snp_value * eff_value

```{r}
join_tbl <- join_tbl %>% 
  mutate(score = snp_value * eff_value)

head(join_tbl)
```

### Column totals

Row totals

```{r}
join_tbl %>% 
  summarize(total_score = sum(score))
```







